Supplemental Materials

### 2021-07-20 supplement/supplement-main.Rmd
made reproducible

Then, we describe our empirical methods in detail. This begins with the qualitative research we conducted with vaccine skeptics to develop psychometrically-valid quantitative scales measuring a network of 14 beliefs related to vaccine skepticism. Then, we report the detailed methods for each of the empirical studies reported in the paper. In Study 1, we used Bayesian network structure learning algorithms to develop a cognitive model of the relationships among these beliefs in a large sample of people. This model can predict how beliefs correlate across people and how they would be revised in the face of evidence. In Study 2, we replicated the success of an existing intervention emphasizing the dangers of childhood diseases (Horne et al., 2015) and used our model to provide a more detailed understanding of _how and why_ this intervention increases vaccine intentions. In Study 3, we drew inspiration from our model to create and validate the success of  a novel intervention aimed at debunking concerns about toxic additives. Finally, in Study 4 we tested whether people's beliefs changed coherently following relevant real-world events, by examining beliefs before and after serious outbreaks of measles in the United States in early 2019.

@@

Figure \@ref(fig:dag-example) shows an example of a simple Bayesian Network. The network itself is a _directed acyclic graph_ (DAG), a type of graphical model. Graphical models represent variables as nodes connected by edges. In a DAG, these edges are directed, flowing from parent to child, and are constrained so that they cannot form cycles or loops. In a Bayesian network, these directed edges encode conditional dependencies among variables, so that the conditional probability distribution for a given variable is defined only in terms of its direct parents. Together, a child node and its parents are called a _family_.

```{r dag-example, echo=F, fig.align='center', out.width = '40%', fig.cap="Example DAG illustrating some variables related to wildfires."}

@@

To conduct an efficient search through the super-exponential space of DAGs, we tested several variants of two structure learning algorithms implemented in the “bnlearn” R package [@scutari2010]. The first is a purely score-based “hill-climbing” algorithm  and the second was a hybrid algorithm that combined the hill-climbing algorithm with a prior “restriction” phase that pruned down the space of DAGs. This pruning was accomplished with the “min-max parents and children” algorithm, which can be tuned with a hyperparameter alpha, which we tested at several levels. To speed computations, we restricted the maximum number of parents in the graph to five for all nodes, as the complexity of the scoring function is exponential in the number of parents. In all resulting networks no node was found to have more than four parents, suggesting this did not impact the final results.

We performed a series of 10-run 10-fold cross validation analyses to evaluate the success of these different structure learning strategies. In addition to the different algorithms, we also compared four different “blacklists” that imposed different constraints on the structure learning process. In addition to our theory-driven blacklist based on our generative model assumptions, we also explored unconstrained structure learning, and two additional blacklists: one which stipulated only that “intentions to vaccinate” should be a child node with no children of its own, and another that stipulated also that the most abstract beliefs “holistic balance” and “naturalism” should be allowed to have no parents other than each other.

Figure \@ref(fig:cross-validation) presents the results of these cross validation tests. As seen in the figure, all structure learning approaches produced models with similar out-of-sample performance, with more variability among folds and runs than between algorithms. Thus the cross-validation does not uniquely identify an optimal structure learning strategy.

@@

```{r cross-validation, echo=FALSE, fig.align='center', fig.height=3, fig.width=4, fig.cap="Cross-validation performance for diffrent structure learning algorithms on the training data from Study 1."}

@@

cv_res %>% 
  ggplot(aes(x=reorder(algo, loss), y=loss)) +
  geom_boxplot(fill="skyblue") +
  coord_flip() +
  labs(y = "Log-likelihood loss", x = "Algorithm") +
  theme_bw() +
  theme(panel.grid = element_blank())

@@

We used a bootstrapping procedure to estimate the confidence in each of the links in this model, conducting the structure learning procedure repeatedly for 200 bootstrapped resamples of the original training data. Figure \@ref(fig:disp-bootgraph) visualizes the frequency with which each arc appeared in the resulting model across these replicates, and figure \@ref(fig:disp-maingraph) visualizes the model with the coefficients for each link encoded by color.

@@

```{r disp-bootgraph, echo=F, fig.width=8, fig.height=5, fig.cap="Bootstrapped confidence estimates for Bayesian network edges. Edge color indicates frequency with which edges appeared when conducting search across 200 bootstrap replicates. Only edges appearing in at least 20\\% of graphs are shown. Our primary outcome of interest, vaccination intentions, is highlighted in yellow."}

@@

```{r disp-maingraph, echo=F, fig.width=8, fig.height=5, fig.align='center', fig.cap="Bayesian network model of intuitive theory surrounding vaccination decisions. Edge color indicates the magnitude of the generative (blue) or preventive (red) coeffient for the links. Our primary outcome of interest, vaccination intentions, is highlighted in yellow."}

@@

We validated our final model by testing its predictions on the held-out testing data. The models’ predictions are plotted against the true values for the testing data in figure X below. As seen in the figure, the quality of predictions varied across beliefs, but in many cases was quite strong.

@@

```{r validation_plot, echo=FALSE, fig.height=6, fig.width=10, fig.align="center", fig.cap="Model predictions versus observed values for test data from Study 1. Dashed lines represent perfect fits."}

@@

plt.valid
```
For each node, we also compared the Bayesian network model’s predictions to a naive “maximal” model using values on all other belief scales as predictors. Our intention was to capture the degree to which each belief is predictable from the others as a whole. As shown in table X, the relatively simpler Bayesian network model does not sacrifice any predictive power compared to the suite of maximal naive models.

Finally, we can also compare the quality of our model predictions to the test-retest reliability of the belief scales (estimated from the pre-test and post-test measures in the control condition of Study 2), establishing an overall upper-bound on the predictability of these measures. As shown in Table \@ref(tab:model_valid_table), the Bayesian network model did approximately as well as possible for seven beliefs, but fell short for the other seven. These failures suggest that the model is missing the information that would be needed to predict these beliefs. Although the notion of an intuitive theory presupposes that there is some collective unit of cognitive machinery defining people’s thinking about a particular domain, defining the boundaries of an intuitive theory is far from straightforward. It could be that new measurements are needed to fully capture these aspects of the intuitive theory, or these beliefs are sufficiently peripheral that they should not be considered part of the intuitive theory. Given the fuzzy nature of the problem, and of cognition more generally, we expect that these types of decisions will be driven as much by investigators' pragmatic concerns as they are by psychological facts.

```{r model_valid_table, echo=FALSE, message=F, warning=F}
# compare S1 model predictions to test-retest reliability?

@@

Returning to the DAG shown in Figure \@ref(fig:dag_example) above, consider the subgraph $FIRE \rightarrow SMOKE$, capturing the fact that fire causes smoke. This relationship is represented as a probabilistic function, where the probability of SMOKE depends on the probability of FIRE. If we observe the presence of SMOKE, we can draw inferences about FIRE.

@@

To model the effects of participants observing evidence, we augmented the Bayesian network representing the cognitive model (shown in Figure \@ref(fig:disp-maingraph)) with an additional node representing the evidence. For instance, to model the effects of the “Disease Risk” intervention (Study 2), we added an evidence node as a child of “disease severity,” acting as virtual evidence for this node. To compute the CPT for this node as a function of disease severity beliefs, we estimated an _evidence ratio_ based on participants’ pretest and posttest credences in disease severity. To so do, we made use of the log-odds expression of Bayes rule,

@@

Figure \@ref(fig:corrplot) presents correlations among the 14 belief scales measured in Study 1. As the figure shows, there are strong correlations among many of the beliefs, suggesting they are related by an underlying intuitive theory.

```{r corrplot, echo=FALSE, message=F, fig.height=7, fig.width=7, fig.align='center', fig.cap = "Correlations among 14 the belief scales measured in Study 1. Correlation matrix is arranged by hierarhical clustering algorithm to highlight connections among beliefs."}
# plot correlation matrix
library(corrplot)

@@

ggplot(temp_cor_df,
       aes(x = reorder(scale1, desc(order1)), 
           y = reorder(scale2, order2), 
           fill = cor)) +
  geom_tile(color = "black") +
  geom_text(aes(label = format(round(cor, 2), digits = 2)), size = 3) +
  scale_fill_distiller(palette ="RdBu", limits = c(-1, 1), direction = 1,
                     guide = guide_colorbar(barheight = 15)) + # direction = -1
  theme_minimal(base_size=12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.title = element_blank(),
        aspect.ratio=1) +
  labs(fill = "Corr.")
```

@@

Procedures for part 1 of Study 2 were nearly identical to Study 1, save for the use of four rather than two attention check questions, and the removal of debriefing statements at the end of the survey. On Day 2 (XX days later), participants completed a study that was designed to be very similar to Horne, Powell, et al.’s (2015) original study. Participants were randomly assigned to either the Disease Risk condition or the No Intervention condition, with the constraint that the distributions of participants’ Day 1 _vaccine intentions_ should be similar across the two conditions; we accomplished this by stratifying participants into groups based on their pretest vaccine intentions, and randomly assigning them to conditions in equal numbers within each group.

Participants were presented with a slightly modified version of the “Disease Risk” intervention from Horne, Powell et al.’s (2015) original study, which was based on materials available on the CDC website (XX). Compared with Horne, Powell et al. (2015), we modified the intervention to present these parts in a fixed (rather than randomized) order, made slight changes to wording to improve clarity, and added a final statement summarizing the take-home message of the intervention. Full text of this intervention can be found in appendix B of these supplementary materials.

@@

Participants in the No Intervention condition (_n_=XX) completed an identical set of questions, except that they did not read any material before completing the 14 belief scales and the questions about their personal background.

@@

```{r s2_model, echo=FALSE, message=F, warning=F}

@@

We conducted a mixed-effects Bayesian ordinal regression to assess the effects of the Disease Risk intervention on responses to the _Vaccination Intentions_ scale items, using the “brms” package for R [@burkner2017]. We regressed responses on phase (Phase 1 vs. 2, dummy-coded with Phase 1 as the baseline), condition (dummy-coded with the Disease Risk condition as the baseline, so as to assess difference scores in our condition of primary interest), and the interaction between phase and condition. We included random intercepts by subject and scale item. In this model, the unique effect of the intervention on vaccination intention responses is assessed by the interaction term (indicating the unique effect of Phase 2 for intervention versus no-intervention participants). The Disease Risk intervention increased participants’ vaccination intention ratings (see Table \@ref(tab:s2_model)). We consider this to be a full replication of Horne et al.’s [-@horne.etal2015] results.

To model the changes in the belief network, we estimate the “evidence ratio” using the method described in the modeling section above. The maximum likelihood estimate of the evidence ratio was XXX. To simulate the impacts of this evidence, we supplied virtual evidence to the _disease severity_ node by adding a child node to the cognitive-model network with a CPT capturing this evidence ratio. Given that our cognitive model is parameterized via noisy-logical functions, the predicted influence of each belief is closely related to its first-order correlation.

Figure \@ref(fig:s2_plot) (right) shows the correlation between predicted and observed belief changes conditional on the Disease Risk intervention.

```{r s2_plot, echo=FALSE}
# create cognitive model for predicting change scores
base_model <- create_bnlearn_cog_model(main_model, train)
```

```{r echo=F, warning=F, message=F, fig.width=8, fig.height=4, fig.cap="Belief changes following the intervention in Experiment 2. Left, mean difference scores from pretest to posttest for the 14 belief scales across experimental conditions; our primary outcome of interest, vaccination intentions, is highlighted in yellow. Right, mean observed and model predicted change scores for the Disease Risk intervention condition. In both plots, error bars represent 95\\% confidence intervals."}
## combined plot for paper
library(patchwork)

@@

cor_val <- format(round(cor(obs_pred$predicted, obs_pred$observed),3), nsmall = 3)

s2_obs_pred_plot <- obs_pred %>% # a bit of a hack
  ggplot(aes(x=predicted, y = observed, ymin = ll, ymax=ul)) +
  geom_abline(slope=1, intercept=0, linetype="dashed", alpha =.5) +
  geom_errorbar( width=0, color="grey") +
  geom_point(size=2.5) +
  annotate(geom="text", x=.03, y=-.04, label= paste("r =", cor_val)) +
  coord_cartesian(xlim=c(-.06,.06), ylim=c(-.06,.06)) +
  theme_bw(base_size=11) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(aspect.ratio = 1) +
  labs(title = "B", y = "Observed change", x="Predicted change")

@@

By assuming that these beliefs are connected via an intuitive theory, we can infer how any given belief in the network will change in response to evidence targeting any other belief in the network. We used the cognitive model to predict how vaccination intentions would be influenced by an intervention targeting each of the 13 other beliefs in the network. Given that our cognitive model is parameterized via noisy-logical functions, the predicted influence of each belief is closely related to its first-order correlation. This captures qualitative Feature 3 of intuitive theories. Figure \@ref(fig:predicted_impacts) presents these predictions, assuming the changes in beliefs are designed to move vaccine intentions in a positive direction (e.g. increasing “disease severity” beliefs but decreasing “vaccine danger” beliefs).

```{r predicted_impacts, echo=F, message=F, fig.align='center', fig.height=3, fig.width=3, fig.cap="Predicted effects on Vaccine Intentions for hypothetical interventions targeting each of the 13 other beliefs"}

@@

```{r, echo=FALSE, message=F, warning=F}
# compare conditions on vaccIntent w/ regression model

@@

```{r, echo=FALSE, message=F, warning=F}

@@

To model the effect of the Vaccine Ingredients intervention, we followed the same procedure as in Study 2, but this time augmented the model with a node representing the Vaccine Ingredients intervention as a child of the _vaccine toxicity_ belief node. The maximum likelihood estimate of the evidence ratio for the Vaccine Ingredients intervention was XXX. To simulate the impacts of this evidence, we supplied virtual evidence to the _vaccine toxicity_ node by adding a child node to the cognitive-model network with a CPT capturing this evidence ratio.

Figure \@ref(fig:s3_plot) (right) shows the correlation between predicted and observed belief changes conditional on the Vaccine Ingredients intervention.

```{r s3_plot, echo=F, message=F, warning=F, fig.width=8, fig.height=4, fig.cap="Belief changes following the intervention in Experiments 3a and 3b. Left, mean difference scores from pretest to posttest for the 14 belief scales across experimental conditions; our primary outcome of interest, vaccination intentions, is highlighted in yellow. Right, mean observed and model predicted change scores for the Vaccine Ingredients intervention condition. In both plots, error bars represent 95\\% confidence intervals."}

@@

cor_val <- format(round(cor(obs_pred$predicted, obs_pred$observed),3), nsmall = 3)

s3_obs_pred_plot <- obs_pred %>% 
  ggplot(aes(x=predicted, y = observed, ymin = ll, ymax=ul)) +
  geom_abline(slope=1, intercept=0, linetype="dashed", alpha =.5) +
  geom_point(size=2.5) +
  geom_errorbar(alpha=.5, width=0) +
  # coord_fixed(ratio=1) +
  coord_cartesian(xlim=c(-.12,.08), ylim=c(-.12,.08)) +
  annotate(geom="text", x=.05, y=-.1, label= paste("r =", cor_val)) +
  theme_bw(base_size=11) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(aspect.ratio = 1) +
  labs(title = "B", y = "Observed change", x="Predicted change")

@@

```{r s3_plot_bystudy, echo=F, message=F, warning=F, fig.width=10, fig.height=5, fig.cap="Belief changes following the intervention in Experiments 3a and 3b, broken out by experiment. Left, mean difference scores from pretest to posttest for the 14 belief scales across experimental conditions; our primary outcome of interest, vaccination intentions, is highlighted in yellow. Right, mean observed and model predicted change scores for the Vaccine Ingredients intervention condition. In both plots, error bars represent 95\\% confidence intervals."}

@@

```{r s4_table, echo=FALSE, message=F, warning=F}
# compare conditions on vaccIntent w/ regression model
# not run yet 5/25/21, 11:52 AM

@@

Mirroring our analyses of interventions, we regressed responses to the _vaccine intentions_ scale items on predictors for phase (pre-outbreak in August and September 2017 vs. post-outbreak May 2019, dummy-coded), news exposure (dummy-coded with non-exposure as the baseline), and an interaction between phase and news exposure onto participants’ scores, using a Bayesian mixed-effects ordinal regression including random intercepts by subject and item. To assess the effects of recent news exposure on _vaccine intentions_, we examined the interaction between phase and news exposure. Participants’ vaccine intentions were increased by recent news exposure compared to their responses 18 months ago (see Table \@ref(tab:s4_table)).

To model the effects of news exposure we followed the same procedure as in previous studies, but here augmented the model with a node representing news exposure as a child of the _vaccine intentions_ node. The maximum likelihood estimate of the evidence ratio for news exposure was XXX. To simulate the impacts of this evidence, we supplied virtual evidence to the _vaccine intentions_ node by adding a child node to the cognitive-model network with a CPT capturing this evidence ratio.

Figure \@ref(fig:s4_plot) (right) shows the correlation between predicted and observed belief changes for participants who were aware of the New York measles outbreak in May 2019.

```{r s4_plot, echo=F, warning=F, message=F, fig.width=8, fig.height=4, fig.cap="Belief changes following the intervention in Study 4. Left, mean difference scores from pretest to posttest for the 14 belief scales among participants who did and did not report exposure to news of the 2019 Measles outbreak in New York; our primary outcome of interest, vaccination intentions, is highlighted in yellow. Right, mean observed and model predicted change scores for those who were exposed to news of the outbreak. In both plots, error bars represent 95\\% confidence intervals."}

@@

cor_val <- format(round(cor(obs_pred$predicted, obs_pred$observed),3), nsmall = 3)

>>>>

Then, we describe our empirical methods in detail. This begins with the qualitative research we conducted with vaccine skeptics to develop psychometrically-valid quantitative scales measuring a network of 14 beliefs related to vaccine skepticism. Then, we report the detailed methods for each of the empirical studies reported in the paper. In Study 1, we used Bayesian network structure learning algorithms to develop a cognitive model of the relationships among these beliefs in a large sample of people. This model can predict how beliefs correlate across people and how they would be revised in the face of evidence. In Study 2, we replicated the success of an existing intervention emphasizing the dangers of childhood diseases [@horne.etal2015] and used our model to provide a more detailed understanding of _how and why_ this intervention increases vaccine intentions. In Study 3, we drew inspiration from our model to create and validate the success of  a novel intervention aimed at debunking concerns about toxic additives. Finally, in Study 4 we tested whether people's beliefs changed coherently following relevant real-world events, by examining beliefs before and after serious outbreaks of measles in the United States in early 2019.

@@

Figure 1 shows an example of a simple Bayesian Network. The network itself is a _directed acyclic graph_ (DAG), a type of graphical model. Graphical models represent variables as nodes connected by edges. In a DAG, these edges are directed, flowing from parent to child, and are constrained so that they cannot form cycles or loops. In a Bayesian network, these directed edges encode conditional dependencies among variables, so that the conditional probability distribution for a given variable is defined only in terms of its direct parents. Together, a child node and its parents are called a _family_.

```{r fig1-dag-example, echo=F, message=F, warning=F, fig.align='center', out.width = '40%', fig.cap="Example DAG illustrating some variables related to wildfires."}

@@

To conduct an efficient search through the super-exponential space of DAGs, we tested several variants of two structure learning algorithms implemented in the “bnlearn” R package [@scutari2010]. The first is a purely score-based “hill-climbing” algorithm(HC) and the second was a hybrid algorithm that combined the hill-climbing algorithm with a prior “restriction” phase that pruned down the space of DAGs (MMHC). This pruning was accomplished with the “min-max parents and children” algorithm, which can be tuned with a hyperparameter alpha, which we tested at several levels. To speed computations, we restricted the maximum number of parents in the graph to five for all nodes, as the complexity of the scoring function is exponential in the number of parents. In all resulting networks no node was found to have more than four parents, suggesting this did not impact the final results.

We performed a series of 10-run 10-fold cross validation analyses to evaluate the success of these different structure learning strategies. In addition to the different algorithms, we also compared four different “blacklists” that imposed different constraints on the structure learning process. In addition to our theory-driven blacklist based on our generative model assumptions ("theory"), we also explored unconstrained structure learning, and two additional blacklists: one which stipulated only that “intentions to vaccinate” should be a child node with no children of its own ("child"), and another that stipulated also that the most abstract beliefs “holistic balance” and “naturalism” should be allowed to have no parents other than each other ("parents").

Figure 2 displays the results of these cross-validation tests. As seen in the figure, all structure learning approaches produced models with similar out-of-sample performance, with more variability among folds and runs than between algorithms. Thus the cross-validation does not uniquely identify an optimal structure learning strategy.

@@

```{r fig2-cross-validation, echo=FALSE, fig.align='center', fig.height=3, fig.width=4, fig.cap="Cross-validation performance for diffrent structure learning algorithms on the training data from Study 1."}

@@

cv_res %>% 
  mutate(algo = case_when(
    algo == "hc" ~ "HC",
    algo == "mmhc_05" ~ "MMHC .05",
    algo == "mmhc_01" ~ "MMHC .01",
    algo == "hc_intentdv" ~ "HC (child)",
    algo == "mmhc_intentdv_05" ~ "MMHC .05 (child)",
    algo == "mmhc_intentdv_01" ~ "MMHC .01 (child)",
    algo == "hc_idv_abspar" ~ "HC (parents)",
    algo == "mmhc_idv_abspar_05" ~ "MMHC .05 (parents)",
    algo == "mmhc_idv_abspar_01" ~ "MMHC .01 (parents)",
    algo == "hc_theory" ~ "HC (theory)",
    algo == "mmhc_theory_05" ~ "MMHC .05 (theory)",
    algo == "mmhc_theory_01" ~ "MMHC .01 (theory)"
  )) %>% 
  ggplot(aes(x=reorder(algo, loss), y=loss)) +
  geom_boxplot(fill="skyblue") +
  coord_flip() +
  labs(y = "Log-likelihood loss", x = "Algorithm") +
  theme_bw() +
  theme(panel.grid = element_blank())

@@

We used a bootstrapping procedure to estimate the confidence in each of the links in this model, conducting the structure learning procedure repeatedly for 200 bootstrapped resamples of the original training data. Figure 3 visualizes the frequency with which each arc appeared in the resulting model across these replicates, and figure \@ref(fig:disp-maingraph) visualizes the model with the coefficients for each link encoded by color.

@@

```{r fig3-disp-bootgraph, echo=F, fig.width=8, fig.height=5, fig.cap="Bootstrapped confidence estimates for Bayesian network edges. Edge color indicates frequency with which edges appeared when conducting search across 200 bootstrap replicates. Only edges appearing in at least 20\\% of graphs are shown. Our primary outcome of interest, vaccination intentions, is highlighted in yellow."}

@@

```{r fig4-disp-maingraph, echo=F, fig.width=8, fig.height=5, fig.align='center', fig.cap="Bayesian network model of intuitive theory surrounding vaccination decisions. Edge color indicates the magnitude of the generative (blue) or preventive (red) coeffient for the links. Our primary outcome of interest, vaccination intentions, is highlighted in yellow."}

@@

We validated our final model by testing its predictions on the held-out testing data. The models’ predictions are plotted against the true values for the testing data in figure 5 below. As seen in the figure, the quality of predictions varied across beliefs, but in many cases was quite strong.

@@

```{r fig5-validation_plot, echo=FALSE, fig.height=6, fig.width=10, fig.align="center", fig.cap="Model predictions versus observed values for test data from Study 1. Dashed lines represent perfect fits."}

@@

plt.valid
```
For each node, we also compared the Bayesian network model’s predictions to a naive “maximal” model using values on all other belief scales as predictors. Our intention was to capture the degree to which each belief is predictable from the others as a whole. As shown in table 1, the relatively simpler Bayesian network model does not sacrifice any predictive power compared to the suite of maximal naive models.

Finally, we can also compare the quality of our model predictions to the test-retest reliability of the belief scales (estimated from the pre-test and post-test measures in the control condition of Study 2), establishing an overall upper-bound on the predictability of these measures. As shown in Table 1, the Bayesian network model did approximately as well as possible for seven beliefs, but fell short for the other seven. These failures suggest that the model is missing the information that would be needed to predict these beliefs. Although the notion of an intuitive theory presupposes that there is some collective unit of cognitive machinery defining people’s thinking about a particular domain, defining the boundaries of an intuitive theory is far from straightforward. It could be that new measurements are needed to fully capture these aspects of the intuitive theory, or these beliefs are sufficiently peripheral that they should not be considered part of the intuitive theory. Given the fuzzy nature of the problem, and of cognition more generally, we expect that these types of decisions will be driven as much by investigators' pragmatic concerns as they are by psychological facts.

```{r tab1-model_valid_table, echo=FALSE, message=F, warning=F}
# compare S1 model predictions to test-retest reliability?

@@

Returning to the DAG shown in Figure 1 above, consider the subgraph $FIRE \rightarrow SMOKE$, capturing the fact that fire causes smoke. This relationship is represented as a probabilistic function, where the probability of SMOKE depends on the probability of FIRE. If we observe the presence of SMOKE, we can draw inferences about FIRE.

@@

To model the effects of participants observing evidence, we augmented the Bayesian network representing the cognitive model (shown in Figure 4) with an additional node representing the evidence. For instance, to model the effects of the “Disease Risk” intervention (Study 2), we added an evidence node as a child of “disease severity,” acting as virtual evidence for this node. To compute the CPT for this node as a function of disease severity beliefs, we estimated an _evidence ratio_ based on participants’ pretest and posttest credences in disease severity. To so do, we made use of the log-odds expression of Bayes rule,

@@

Figure 6 presents correlations among the 14 belief scales measured in Study 1. As the figure shows, there are strong correlations among many of the beliefs, suggesting they are related by an underlying intuitive theory.

```{r fig6-corrplot, echo=FALSE, message=F, fig.height=7, fig.width=7, fig.align='center', fig.cap = "Correlations among 14 the belief scales measured in Study 1. Correlation matrix is arranged by hierarhical clustering algorithm to highlight connections among beliefs."}
# plot correlation matrix
library(corrplot)

@@

ggplot(temp_cor_df,
       aes(x = reorder(scale1, desc(order1)), 
           y = reorder(scale2, order2), 
           fill = cor)) +
  geom_tile(color = "black") +
  geom_text(aes(label = format(round(cor, 2), digits = 2)), size = 2.75) +
  scale_fill_distiller(palette ="RdBu", limits = c(-1, 1), direction = 1,
                     guide = guide_colorbar(barheight = 15)) + # direction = -1
  theme_minimal(base_size=12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.title = element_blank(),
        aspect.ratio=1) +
  labs(fill = "Corr.")
```

@@

Procedures for part 1 of Study 2 were nearly identical to Study 1, save for the use of four rather than two attention check questions, and the removal of debriefing statements at the end of the survey. On Day 2 (approximately two days later), participants completed a study that was designed to be very similar to Horne, Powell, et al.’s [-@horne.etal2015] original study. Participants were randomly assigned to either the Disease Risk condition or the No Intervention condition, with the constraint that the distributions of participants’ Day 1 _vaccine intentions_ should be similar across the two conditions; we accomplished this by stratifying participants into groups based on their pretest vaccine intentions, and randomly assigning them to conditions in equal numbers within each group.

Participants were presented with a slightly modified version of the “Disease Risk” intervention from Horne et al.’s [-@horne.etal2015] original study, which was based on materials available on the CDC website. Compared with Horne et al. [-@horne.etal2015], we modified the intervention to present these parts in a fixed (rather than randomized) order, made slight changes to wording to improve clarity, and added a final statement summarizing the take-home message of the intervention. Full text of this intervention can be found in appendix B of these supplementary materials.

@@

Participants in the No Intervention condition (_n_ = `r sum(s2_demo$condition=="noInterv")`) completed an identical set of questions, except that they did not read any material before completing the 14 belief scales and the questions about their personal background.

@@

```{r tab2-s2_model, echo=FALSE, message=F, warning=F}

@@

We conducted a mixed-effects Bayesian ordinal regression to assess the effects of the Disease Risk intervention on responses to the _Vaccination Intentions_ scale items, using the “brms” package for R [@burkner2017]. We regressed responses on phase (Phase 1 vs. 2, dummy-coded with Phase 1 as the baseline), condition (dummy-coded with the Disease Risk condition as the baseline, so as to assess difference scores in our condition of primary interest), and the interaction between phase and condition. We included random intercepts by subject and scale item. In this model, the unique effect of the intervention on vaccination intention responses is assessed by the interaction term (indicating the unique effect of Phase 2 for intervention versus no-intervention participants). The Disease Risk intervention increased participants’ vaccination intention ratings (see Table 2). We consider this to be a full replication of Horne et al.’s [-@horne.etal2015] results.

To model the changes in the belief network, we estimate the “evidence ratio” using the method described in the modeling section above. To simulate the impacts of this evidence, we supplied virtual evidence to the _disease severity_ node by adding a child node to the cognitive-model network with a CPT capturing this evidence ratio. Given that our cognitive model is parameterized via noisy-logical functions, the predicted influence of each belief is closely related to its first-order correlation.

Figure 7 (right) shows the correlation between predicted and observed belief changes conditional on the Disease Risk intervention.

```{r, echo=FALSE}
# create cognitive model for predicting change scores
base_model <- create_bnlearn_cog_model(main_model, train)
```

```{r fig7-s2_plot, echo=F, warning=F, message=F, fig.width=8, fig.height=4, fig.cap="Belief changes following the intervention in Experiment 2. Left, mean difference scores from pretest to posttest for the 14 belief scales across experimental conditions; our primary outcome of interest, vaccination intentions, is highlighted in yellow. Right, mean observed and model predicted change scores for the Disease Risk intervention condition. In both plots, error bars represent 95\\% confidence intervals."}
## combined plot for paper
library(patchwork)

@@

cor_val_s2 <- format(round(cor(obs_pred$predicted, obs_pred$observed),3), nsmall = 3)

s2_obs_pred_plot <- obs_pred %>% # a bit of a hack
  ggplot(aes(x=predicted, y = observed, ymin = ll, ymax=ul)) +
  geom_abline(slope=1, intercept=0, linetype="dashed", alpha =.5) +
  geom_errorbar( width=0, color="grey") +
  geom_point(size=2.5) +
  annotate(geom="text", x=.03, y=-.04, label= paste("r =", cor_val_s2)) +
  coord_cartesian(xlim=c(-.06,.06), ylim=c(-.06,.06)) +
  theme_bw(base_size=11) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(aspect.ratio = 1) +
  labs(title = "B", y = "Observed change", x="Predicted change")

@@

By assuming that these beliefs are connected via an intuitive theory, we can infer how any given belief in the network will change in response to evidence targeting any other belief in the network. We used the cognitive model to predict how vaccination intentions would be influenced by an intervention targeting each of the 13 other beliefs in the network. Given that our cognitive model is parameterized via noisy-logical functions, the predicted influence of each belief is closely related to its first-order correlation. This captures qualitative Feature 3 of intuitive theories. Figure 8 presents these predictions, assuming the changes in beliefs are designed to move vaccine intentions in a positive direction (e.g. increasing “disease severity” beliefs but decreasing “vaccine danger” beliefs).

```{r fig8-predicted_impacts, echo=F, message=F, fig.align='center', fig.height=3, fig.width=3, fig.cap="Predicted effects on Vaccine Intentions for hypothetical interventions targeting each of the 13 other beliefs"}

@@

```{r tab3-s3_reg, echo=FALSE, message=F, warning=F}
# compare conditions on vaccIntent w/ regression model

@@

```{r tab4-tab5, echo=FALSE, message=F, warning=F}

@@

To model the effect of the Vaccine Ingredients intervention, we followed the same procedure as in Study 2, but this time augmented the model with a node representing the Vaccine Ingredients intervention as a child of the _vaccine toxicity_ belief node.  To simulate the impacts of this evidence, we supplied virtual evidence to the _vaccine toxicity_ node by adding a child node to the cognitive-model network with a CPT capturing this evidence ratio.

Figure 9 (right) shows the correlation between predicted and observed belief changes conditional on the Vaccine Ingredients intervention.

```{r fig9-s3_plot, echo=F, message=F, warning=F, fig.width=8, fig.height=4, fig.cap="Belief changes following the intervention in Experiments 3a and 3b. Left, mean difference scores from pretest to posttest for the 14 belief scales across experimental conditions; our primary outcome of interest, vaccination intentions, is highlighted in yellow. Right, mean observed and model predicted change scores for the Vaccine Ingredients intervention condition. In both plots, error bars represent 95\\% confidence intervals."}

@@

cor_val_s3 <- format(round(cor(obs_pred$predicted, obs_pred$observed),3), nsmall = 3)

s3_obs_pred_plot <- obs_pred %>% 
  ggplot(aes(x=predicted, y = observed, ymin = ll, ymax=ul)) +
  geom_abline(slope=1, intercept=0, linetype="dashed", alpha =.5) +
  geom_point(size=2.5) +
  geom_errorbar(alpha=.5, width=0) +
  # coord_fixed(ratio=1) +
  coord_cartesian(xlim=c(-.12,.08), ylim=c(-.12,.08)) +
  annotate(geom="text", x=.05, y=-.1, label= paste("r =", cor_val_s3)) +
  theme_bw(base_size=11) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(aspect.ratio = 1) +
  labs(title = "B", y = "Observed change", x="Predicted change")

@@

```{r fig10-s3_plot_bystudy, echo=F, message=F, warning=F, fig.width=10, fig.height=5, fig.cap="Belief changes following the intervention in Experiments 3a and 3b, broken out by experiment. Left, mean difference scores from pretest to posttest for the 14 belief scales across experimental conditions; our primary outcome of interest, vaccination intentions, is highlighted in yellow. Right, mean observed and model predicted change scores for the Vaccine Ingredients intervention condition. In both plots, error bars represent 95\\% confidence intervals."}

@@

```{r tab6-s4_table, echo=FALSE, message=F, warning=F}
# compare conditions on vaccIntent w/ regression model
# not run yet 5/25/21, 11:52 AM

@@

Mirroring our analyses of interventions, we regressed responses to the _vaccine intentions_ scale items on predictors for phase (pre-outbreak in August and September 2017 vs. post-outbreak May 2019, dummy-coded), news exposure (dummy-coded with non-exposure as the baseline), and an interaction between phase and news exposure onto participants’ scores, using a Bayesian mixed-effects ordinal regression including random intercepts by subject and item. To assess the effects of recent news exposure on _vaccine intentions_, we examined the interaction between phase and news exposure. Participants’ vaccine intentions were increased by recent news exposure compared to their responses 18 months ago (see Table 6).

To model the effects of news exposure we followed the same procedure as in previous studies, but here augmented the model with a node representing news exposure as a child of the _vaccine intentions_ node. To simulate the impacts of this evidence, we supplied virtual evidence to the _vaccine intentions_ node by adding a child node to the cognitive-model network with a CPT capturing this evidence ratio.

Figure 11 (right) shows the correlation between predicted and observed belief changes for participants who were aware of the New York measles outbreak in May 2019.

```{r fig11-s4_plot, echo=F, warning=F, message=F, fig.width=8, fig.height=4, fig.cap="Belief changes following the intervention in Study 4. Left, mean difference scores from pretest to posttest for the 14 belief scales among participants who did and did not report exposure to news of the 2019 Measles outbreak in New York; our primary outcome of interest, vaccination intentions, is highlighted in yellow. Right, mean observed and model predicted change scores for those who were exposed to news of the outbreak. In both plots, error bars represent 95\\% confidence intervals."}

@@

cor_val_s4 <- format(round(cor(obs_pred$predicted, obs_pred$observed),3), nsmall = 3)
